ACT Cert FMM Template

Study unit 2 : modelling the maths of debt

present values, future cash flows, timing and risk, interest rate sensitivity and duration models.

Value relationships with respect to yield maturity coupon rate and coupon frequency.

1 The idea of interest

Interest may be regarded as a reward paid by one person or organisation (the borrower) for the use of an asset, referred to as capital, belonging to another person or organisation (the lender). Capital and interest need to be measured in terms of the same commodity, but when expressed in monetary terms, capital is also referred to as principal.

If there is some risk of default (i.e. loss of capital or non-payment of interest) a lender would expect to be paid a higher rate of interest than would otherwise be the case.
Another factor which may influence the rate of interest on any transaction is an allowance for the possible depreciation or appreciation in the value of the currency in which the transaction is carried out. This factor is obviously very important in times of high inflation. The most elementary concept is that of simple interest. This leads naturally to the idea of compound interest, which is much more commonly found in practice — at least in relation to all but short-term investments. Both concepts are easily described within the
framework of a savings account.
 
 

%===========================================================================================

Define an equation of value.
1. Define an equation of value, where payment or receipt is certain.
2. Describe how an equation of value can be adjusted to allow for uncertain receipts or payments.
3. Understand the two conditions required for there to be an exact solution to an equation of value.
%===============================================================================%


%Cert FM Section 10

Describe the investment and risk characteristics of the following types of asset available for investment purposes: fixed interest government borrowings fixed interest borrowing by other bodies shares and other equity-type finance derivatives
