1.C.4.

Covered Interest Arbitrage

Offices
-Front office
-Middle office
-Back office

1.C.5.

(k) 
Define Listing and Float

1.C.6


1.C.7.

(a) 
Commodities are divided into 4 types
- Metals
- Softs : Cocoa, Coffee and Rubber
- Grains and Oilseeds : Most edible agricultural products
- Livestock : Live animals and meat products
(b)
Metals are defined as

1) Base - non ferrous metals.
2) Strategic  - Bismuth and Vanadium
3) Minor - Cobalt and Chromium
4) Precious - Gold Silver and Palladium

(c)
Contrast grains, Oilseeds, and fibers

Grains : Wheat Barley Rice and Oats.

Oilseeds : Soybeans, Palm Kernel and Flaxseed.

Fibers : Wood Cloths and Silks.


(d)
Define "on the spot" and "settlement of difference"

Spot contracts are settled on the spot.

(e) 
Free Alongside Ship (FAS)
An Incoterm (FAS) that means the seller is responsible for the cost of transporting and delivering goods alongside a vessel
in a port in his or her country. Since the buyer has responsibility for export clearance under FAS, it is not a practical 
Incoterm for U.S. exports. FAS should be used only for ocean shipments since risk and responsibility shift from seller to 
buyer when the goods are placed within the reach of the ship's tackle (crane).

Convenience Yield Theory


1.C.8.

Price volatility is caused by fundamental factors such as supply/Demand, and weather and financial factors
such as technical trading, speculators and market imperfections.


-----------------------------------------------------------------------------------------------------------
