APRM Section H  Case Studies
 



%=============================================================%
Bankers Trust
Bankers Trust was sued by four of its major clients Federal Paper Board Company, Gibson Greetings, Air Products and Chemical, and Procter & Gamble.
Bankers Trust and P & G reached an out of court settlement regarding all their disputes on derivatives contracts.


Aggressive sales practices, complex product structure, inadequate disclosures – all of these led to problems between Bankers Trust and its important clients.

%=============================================================%

Barings Bank

%=============================================================%
Bankgesellschaft Berlin
Management of Bankgesellschaft Berlin comprised prominent politicians of the Christian Democratic party
BgB planned to acquire strategic banking assets in the Polish market

%=============================================================%
Continental Illinois
Credit Lyonnais
The 8 million depositors did not suffer from any loss, but the French taxpayer took a big hit.
Foreign Currency risk was not a factor.

%=============================================================%
Daiwa Bank
Daiwa bank lost $1.1 billion because of the dealings by Toshihide Iguchi.
About $377 million of Daiwa’s customers’ securities and about $733 million of Daiwa’s own investment securities had been sold off by Iguchi to cover his trading losses.

%=============================================================%
LTCM
The strategy of getting creditors to lend money and invest equity in LTCM had the effect of reducing leverage
LTCM'S balance sheet as of August 31,1998 showed the following: $125 billion in assets,$2.3billion in equity.

%=============================================================%
Metallgesellschaft
MG absorbed the following risks in so far as the structure of MG's deals would create losses if the futures price curves of each of the commodities shifted from mostly backwardation to mostly contango.

        I. Crude commodity price risk
        II. Gasoline commodity price risk 
        III. Heating oil commodity price risk
        IV. Price structure risk
 
National Australian Bank - FX

%=============================================================% 
\subsection{Orange County}
Hike in interest rates by Federal reserve was the main reason for the fall in value of the investment pool.

\subsection{Riggs Bank }
\subsubsection{Equatorial Guinea}
By 2003, the E.G. accounts represented the largest relationship at Riggs Bank, with aggregate deposits ranging from $400 to $700 million at a time.
 

General Augusto Pinochet held an account at Riggs Bank.
Equatorial Guinea's government was accused of embezzling state funds from their account at Riggs Bank.
Riggs Bank was fined $25 million for violations of money-laundering laws.
The bank and Allbritton family agreed to pay $9 million to Pinochet victims for concealing and illegally facilitating movement of Pinochet money out of Britain.

Riggs did produce three KYC client profiles prepared during 1998, 1999, and 2002. Riggs did not produce any KYC documentation related to the opening of this account in December 1994 in the United States.

%=============================================================%

\subsection{Worldcom}
When accounting for an acquisition, WorldCom would write down in one quarter millions of dollars of assets it acquired, while at the same time it "included in this charge against earnings the cost of company expenses expected in the future. The result was bigger losses in the current quarter but smaller ones in future quarters, so that its profit picture would seem to be improving.”  
 
With MCI acquisitions, it reduced the book value of some of the assets of MCI by several billion dollars, and increased the value of "good will," that is, intangible assets-a brand name, by the same amount.
This enabled WorldCom each year to charge a smaller amount against earnings by spreading these large expenses over decades rather than years. The net result was WorldCom's ability to cut annual expenses, acknowledge all MCI revenue and boost profits from the acquisition.  
 
 On its accounts receivables, reserves for bad debts were not made. 
 

The company founder is Bernie Ebbers.
 
From its humble beginnings as an obscure long distance telephone company WorldCom, through the execution of an aggressive acquisition strategy, evolved into the second-largest long distance telephone company in the United States and one of the largest companies handling worldwide Internet data traffic.
 
Arthur Anderson was the company auditor.
"Auditors and analysts are participants in a game of nods and winks." Arthur Levitt Jr., the then chairman of SEC
 

\subsubsection{Bernie Ebbers}
 
The $341 million loan the board granted Mr. Ebbers is the largest amount any publicly traded company has lent to one of its officers in recent memory.  
 
WorldCom charged Ebbers slightly more than 2percent interest, a rate considerably below average.
 
 Bernie Ebbers purchased WorldCom stock on margin 

