PRMIA 4B By-Laws
Describe Article 1: Name, Corporate Offices And Official Language
Describe Article 2: Mission Statement And Purpose
Describe Article 3: Membership
Describe Article 4: Meetings Of Members And Regional Directors
Describe Article 5: Board Of Directors
Describe Article 6: Officers And Duties
Describe Article 7: Committees
Describe Article 8: Regional Chapters
Describe Article 9: Finances
Describe Article 10: Indemnification
Describe Article 11: General Provisions
Describe Article 12: Amendments 

Article 1
 
Article 2
 
Article 3
 
Article 4
 
%==========================================================================%


\section{PRMIA ByLaws}
\subsection{Articles}
\subsubsection{Article 1- Name, Corporate Offices And Official
Language}Offices as determined by B.O.D. \\No Corporate Seal.\\
Language is English.
\subsubsection{Article 2- Mission Statement And Purpose}
\subsubsection{Article 3- Membership} Regular
membership for individual members\\ A Corporation, association,
educational institution or government agency may apply to be an
affiliate member\\ No Transfer of membership.\\Resignation
procedure - advise the membership secretary.
\\Termination of membership may be carried out with 3/4 vote of
B.O.D.
\subsubsection{Article 4 - Meetings Of Members And Regional
Directors} \begin{itemize}\item Locations as determined by
B.O.D.\item Annual Meeting as determined by B.O.D. \item Special
Meetings may be convened following vote by B.O.D. or by 500
petitions by members (at least 10\% from each region). \item
Remote Communication \item Notice of Meeting \item Quorum - at
least 6\% of membership - comprised equally from each region (2\%
each) present in person or by proxy.\item Meetings subject to
Delware General Corporation Law \item Minutes of any meeting of
the Members shall be posted to the PRMIA web site within 30 days
of the conclusion of the meeting.
\end{itemize}
\subsubsection{Article 5 - Board Of Directors}
\begin{itemize}
\item 7 on Board Of Directors \item 4 groups of nomination.
\\- American Hemisphere(2 Directors)
\\- EMEA(2 Directors)
\\- Asia Pacific (2 Directors)
\\- Any Region (3 Directors)
\item Quorum and Voting - 4 must be present. Each has one vote.
Simple Majority of those present.
\end{itemize}
\subsubsection{Article 6 - Officers And Duties}
The officers of PRMIA shall be a Chair, a Vice-Chair, a Secretary
and a Treasurer.\\Other officers, with powers and duties not
inconsistent with these Bylaws as the Board of Directors may from
time to time elect. \\The same person may hold more than one
position provided that the Chair shall not serve as Vice-Chair or
Secretary. \\The Chair, Vice-Chair and Secretary must be
Directors. All other Officers need not be Directors.

\subsubsection{Article 7- Committees}
\begin{itemize} \item An Executive Committee\item A Finance Committee
\item An Education and Certification Committee \item  A Regional
Directors Committee.
\end{itemize}
\subsubsection{Article 8- Regional Chapters}
\subsubsection{Article 9- Finances}
\begin{itemize} \item Deposits and Borrowings as usual. no
provision for loans. \item Ownership of assets by corporation.
Sale of assets - small by 2/3 BOD - large by vote.
\end{itemize}
\subsubsection{Article 10- Indemnification}

\subsubsection{Article 11- General Provisions}
\begin{itemize} \item Contributions\item Waivers - may be deemed
equivalent to notice \item Entering into contracts - officers may
do so with approval of B.O.D. \item Electronic Transmissions.
\end{itemize}

\subsubsection{Article 12- Amendments}
